---
title: "Housing_Price"
format: html
---
```{r}
data = read.csv("D:/Fall/Predictive/Housing Data/data.csv")
data
```
```{r}
print(paste("Number of Rows: ", nrow(data)))
print(paste("Number of Columns: ", ncol(data)))

```
```{r}
summary(data)
```
```{r}
unique(data$city)
```
```{r}
data_df = data[c("price","bedrooms","bathrooms","sqft_living","sqft_lot","floors","condition","view","yr_built")]
data_df
```
```{r}
sum(is.na(data_df)) #Checking NUll Values
```
```{r}
data_df$yr_old = as.integer(format(Sys.Date(),"%Y"))- data_df$yr_built
data_df
```
```{r}
drops = c("yr_built","oldbuilt")

library(dplR)
data_df = data_df[,setdiff(names(data_df),"yr_built","oldbuilt")

data_df = data_df[,!(names(data_df)%in%drops)]
                  
data_df

```
```{r}
cor(data_df)
```
```{r}
library(ggcorrplot)

corr = round(cor(data_df),1)

ggcorrplot(corr,
           type = "lower",
           lab = TRUE,
           lab_size = 3,
           colors= c("indianred1","white", "lightgreen"),
           title = "Correlation Chart",
           ggtheme = theme_light())
```
```{r}
## Scatterplot Matrix 

pairs(~ bedrooms + sqft_living + floors + condition, data = data_df,
      main = "Scatterplot Matrix")

```
#There appears to be a trend in this cell, where properties with more bedrooms tend to have larger square footage (sqft_living). This indicates a positive correlation.

# No clear trend or pattern is visible, suggesting that the condition of the property does not vary systematically with the number of bedrooms.

#This also lacks a clear trend, implying little or no direct relationship between the number of floors and the condition rating.


```{r}
par(mfrow =c(2,3))
boxplot(data_df$bedrooms,main ="Bedrooms")
boxplot(data_df$sqft_living,main ="Sqft_Living")
boxplot(data_df$floors,main ="Floors")
boxplot(data_df$condition,main ="Condition")
boxplot(data_df$view,main ="View")
boxplot(data_df$yr_old,main ="Yr_old")
```
# Sqft_Living and Bedrooms have notable outliers, indicating a few properties with significantly higher values compared to the norm.

#Variables like Sqft_Living show skewed distributions, while others like Condition and Floors are more evenly distributed.

#Variables such as Floors and Condition are tightly distributed, suggesting limited variation in the number of floors and condition ratings across properties.



```{r}
theme_set(theme_classic())

plot = ggplot(data_df,aes(bedrooms,floors))

plot + geom_count(col="tomato3",show.legend = T) +
  labs(y="Floors",
       x="Bedrooms",
       title = "Bedroom vs Floors")

```
The chart shows that most properties in the dataset have 2 to 4 bedrooms and 1 to 2 floors, suggesting a predominance of standard family homes. 

The most common configuration is 2 bedrooms with 2 floors. Larger homes with more bedrooms or floors are relatively rare, indicating that these might be luxury or custom properties.

```{r}
plot(x =data_df$sqft_living,y= data_df$sqft_lot,
     xlab = "sqft_living",
     ylab = "sqft_lot",
     xlim = c(0,3000),
     ylim = c(0,20000),
     main = "sqft_living Vs sqft_lot")
```
There is no clear linear correlation, with a wide spread of data points indicating variability. Most properties cluster around lower square footage values, while a few outliers exhibit larger sizes in both categories. This suggests that additional factors may influence the relationship between these two variables.

```{r}



#detach("package:e1071", unload = TRUE)  # Unload the e1071 package

#install.packages("e1071")


library(e1071)

par(mfrow =c(2,3))

plot(density(data_df$bedrooms),main ="Density Plot: Bedrooms", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$bedrooms),2)))
polygon(density(data_df$bedrooms),col = "green")

plot(density(data_df$sqft_living),main ="Density Plot: Sqft_living", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$sqft_living),2)))
polygon(density(data_df$sqft_living),col ="orange")


plot(density(data_df$sqft_lot),main ="Density Plot: Sqft_lot", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$sqft_lot),2)))
polygon(density(data_df$sqft_lot),col ="green")



plot(density(data_df$condition),main ="Density Plot: Condition", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$condition),2)))
polygon(density(data_df$condition),col ="yellow")

plot(density(data_df$floors),main ="Density Plot: Floors", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$floors),2)))
polygon(density(data_df$floors),col ="pink")

plot(density(data_df$yr_old),main ="Density Plot: Yr_old", ylab ="Frequency",
     sub= paste("Skewness:",round(e1071 ::skewness(data_df$yr_old),2)))
polygon(density(data_df$yr_old),col ="purple")











```
```{r}

# Regression between Sqft_living and Price

ggplot(data_df,aes(y=price,x=sqft_living)) +
  geom_point() +
  xlim(0,9000)+
  ylim(0,5000000)+
  geom_smooth(formula = y~x,method ="lm")


```
```{r}

model = lm(price~bedrooms + sqft_living+ floors+ sqft_lot+condition + view + yr_old, data = data_df)

summary(model)

```
The model suggests that sqft_living, floors, condition, view, and yr_old have a significant relationship with price, while bedrooms and sqft_lot also influence the price, but their effect may be less intuitive.

A key takeaway is that larger homes (sqft_living), better condition, and properties with a view tend to be more expensive, while older homes and homes with more bedrooms may decrease the price.

```{r}
````






